---
title: "data_exploration_econ" 
author: "Katie Yamabe"
format: docx
editor: visual
---

## Libraries

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

LOAD the libraries and the data's

```{r}
library(tidyverse)
library(fixest)
library(rio)
library(lubridate)
library(ggplot2)
```

Load Google Trends Data

```{r}
trends_up_to_list <- list.files(pattern = "trends_up_to_", full.names = TRUE)
trends_up_to <-import_list(trends_up_to_list,rbind = TRUE, fill = TRUE)

```

Aggregating the Google Trends

```{r}
trends_up_to <- trends_up_to %>%   
  mutate(temp = ymd(str_sub(monthorweek,end = 10)))%>%    
  mutate(first_of_month = floor_date(temp, unit = "month"))

trends_up_to <- trends_up_to %>%    
  group_by(schname, keyword)%>%   
  mutate(si = ((index - mean(index))/sd(index)),na.rm = TRUE)
```

Loading the Score Card

```{r}
score <- import("Most+Recent+Cohorts+(Scorecard+Elements).csv")
score <- score %>%    
  mutate(opeid = OPEID)%>%
  filter(score$PREDDEG == 3)

id_name_link <- import("id_name_link.csv") 

id_name_link <- id_name_link %>%    
  group_by(schname) %>%    
  mutate (n = n())%>%    
  filter(n == 1) %>%
  ungroup()
 


```

Combine the Google Trends and Score Card

```{r}
id_name_link <- inner_join(trends_up_to, id_name_link, by = "schname")
 
clean_data <- inner_join(id_name_link, score, by = c("unitid" = "UNITID", "opeid" = "OPEID"))

 

```

Removing Na's and determining that "high income" is the top 25% of the population. According to the 2022 US Census the top 25% income was \$94,001 for the average household. <https://www.census.gov/content/dam/Census/library/publications/2023/demo/p60-279.pdf> (pg7).

```{r}
clean_data <- clean_data %>% 
  mutate(md_earn_wne_p10_REPORTED_EARNINGS = as.character(`md_earn_wne_p10-REPORTED-EARNINGS`)) %>%
  mutate(md_earn_wne_p10_REPORTED_EARNINGS = na_if(md_earn_wne_p10_REPORTED_EARNINGS, "PrivacySuppressed")) %>%
  mutate(md_earn_wne_p10_REPORTED_EARNINGS = ifelse(md_earn_wne_p10_REPORTED_EARNINGS %in% c("", "NULL"), NA, md_earn_wne_p10_REPORTED_EARNINGS)) %>%
  mutate(md_earn_wne_p10_REPORTED_EARNINGS = as.numeric(md_earn_wne_p10_REPORTED_EARNINGS))

clean_data <- clean_data %>%
  mutate(high_earning = ifelse(!is.na(`md_earn_wne_p10-REPORTED-EARNINGS`) & `md_earn_wne_p10-REPORTED-EARNINGS` >= 94001, 1, 0))

clean_data <- clean_data %>%
   select(schname, first_of_month, si, PREDDEG, high_earning, `md_earn_wne_p10-REPORTED-EARNINGS`)

clean_data <- clean_data %>%
  mutate(post_score = ifelse(first_of_month >= as.Date("2015-09-01"),1,0)) %>%
  mutate(med_earn = `md_earn_wne_p10-REPORTED-EARNINGS`)

#regression

reg1 <- lm(si ~ high_earning * post_score, data = clean_data)
summary(reg1)

# Create the plot
ggplot(clean_data, aes(x = first_of_month, y = si, color = factor(high_earning))) +
  geom_line() +
  labs(title = "Search Interest Over Time for High vs. Low Earning Colleges",
       x = "Date",
       y = "Average Z Score",
       color = "Earnings Category") +
  facet_wrap(~ post_score)
```

Reason for Analysis

The analysis conducted was due to filtering down the information to the high-earning colleges as well as pre and post score card. These were key components of the research question. The regression that was ran was the standard index regressed on the high earning colleges after the scorecard came out.

Conclusion

The P-value is statistically significant to the 0.001 significance level. The introduction of the College Scorecard increased search activity on Google Trends for colleges with high-earning graduates by 17.70 percentage points relative to what it did for colleges with low-earning graduates with a standard error of 0.005379. This result comes from the high_earning coefficients in my regression.
